---
title: 'Bios 6301: Assignment 7'
author: "Jeongwon Choi"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

*Due Thursday, 03 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework7.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework7.rmd` or include author name may result in 5 points taken off.

```{r}
library(tidyverse)
```

### Question 1 ###

**21 points**

Use the following code to generate data for patients with repeated measures of A1C (a test for levels of blood glucose).

```{r}
genData <- function(n) {
    if(exists(".Random.seed", envir = .GlobalEnv)) {
        save.seed <- get(".Random.seed", envir= .GlobalEnv)
        on.exit(assign(".Random.seed", save.seed, envir = .GlobalEnv))
    } else {
        on.exit(rm(".Random.seed", envir = .GlobalEnv))
    }
    set.seed(n)
    subj <- ceiling(n / 10)
    id <- sample(subj, n, replace=TRUE)
    times <- as.integer(difftime(as.POSIXct("2005-01-01"), as.POSIXct("2000-01-01"), units='secs'))
    dt <- as.POSIXct(sample(times, n), origin='2000-01-01')
    mu <- runif(subj, 4, 10)
    a1c <- unsplit(mapply(rnorm, tabulate(id), mu, SIMPLIFY=FALSE), id)
    data.frame(id, dt, a1c)
}
x <- genData(500)
```

Perform the following manipulations: (3 points each)

1. Order the data set by `id` and `dt`.
```{r}
x <- x[order(x$id, x$dt),]
rownames(x) <- NULL
head(x)
```

2. For each `id`, determine if there is more than a one year gap in between observations.  Add a new row at the one year mark, with the `a1c` value set to missing.  A two year gap would require two new rows, and so forth.
```{r}

# making a loop that calculates the number of days between the lab test
x$gap <- 0
for (i in 2:nrow(x)) {
  if(x$id[i-1]==x$id[i]){
  x$gap[[i]]=x$dt[i]-x$dt[i-1]
  }
}

# mark refers to the year between the lab tests
x$mark <- x$gap %/% 365
head(x)

#assigning the one year mark to a new row
library(tidyverse)
x$idx <- 1:nrow(x)

head(x)
nr <-nrow(x)

for (i in 1:nr){
 this_idx <- which(x$idx==i)
 this_mark <- x[this_idx, "mark"]
 if(this_mark==0 ){
   next} else{
for (j in 1:this_mark){
  
  new.dt <- x[this_idx-1,'dt']+as.difftime(365,units="days")
  x <- x %>% add_row(id=x[this_idx, "id"], dt=new.dt, a1c=NA, .before=this_idx)
  # x[this_idx-1,'gap'] <- x$dt[this_idx-1]-x$dt[this_idx-2]
  x$gap[[this_idx]] =  x$dt[this_idx]-x$dt[this_idx-1]
}
}
}
x


```

3. Create a new column `visit`.  For each `id`, add the visit number.  This should be 1 to `n` where `n` is the number of observations for an individual.  This should include the observations created with missing a1c values.
```{r}
library(tidyverse)
x$visit <- NULL
uniqueid <- unique(x$id)
for (i in uniqueid) {
  this_id <- which(x$id==i)
  count_this_id <- length(this_id)
  x[this_id,'visit'] <- 1:count_this_id
}
```

4. For each `id`, replace missing values with the mean `a1c` value for that individual.
```{r}
#create a mean a1c value
mean_a1c <- x %>%
group_by(id) %>%
summarize(Mean = mean(a1c, na.rm=TRUE))
mean_a1c

#replace NA with the mean
uniqueid2 <- unique(x$id)
missingidx <- is.na(x$a1c)

for(i in uniqueid2) {
  this_id2 <- x$id==i
  both <- this_id2 & missingidx
  x[both,'a1c'] <- mean_a1c[i,'Mean']
}
```

5. Print mean `a1c` for each `id`.
```{r}
x %>%
group_by(id) %>%
summarize(Mean = mean(a1c))
mean_a1c
```

6. Print total number of visits for each `id`.
```{r}
lastobs <- x %>%
  group_by(id) %>%
  summarise_all(last)

lastobs[,c(1,7)]
```

7. Print the observations for `id = 15`.
```{r}
x[x$id == 15,]
```



### Question 2 ###

**16 points**

Install the `lexicon` package. Load the `sw_fry_1000` vector, which contains 1,000 common words.

```{r}
library(lexicon)
data('sw_fry_1000', package = 'lexicon')
head(sw_fry_1000)
```

1. Remove all non-alphabetical characters and make all characters lowercase. Save the result as `a`.

```{r}
a1 <- tolower(sw_fry_1000)
a <- gsub("[^a-z]", "", a1)
```

Use vector `a` for the following questions. (2 points each)

2. How many words contain the string "ar"?

```{r}
length(grep("ar", a, value=TRUE))
```
64 words contain the string "ar". 

3. Find a six-letter word that starts with "l" and ends with "r".

```{r}
b1 <- grep("^l", a, value=TRUE)
b2 <- grep("r$", b1, value=TRUE)
str_length(b2)
```
letter starts with "l" and ends with "r".

4. Return all words that start with "col" or end with "eck".

```{r}
grep("^col|eck$",a, value=TRUE)
```

5. Find the number of words that contain 4 or more adjacent consonants. Assume "y" is always a consonant.

```{r}
length(grep("[^aeiou]{4}", a, value=TRUE))
```
8 words contain 4 or more adjacent consonants


6. Return all words with a "q" that isn't followed by a "ui".

```{r}
c1 <- grep("[q]", a, value=TRUE)
idx <- grep("[q](!?ui)", c1)
c1[-idx]
```

7. Find all words that contain a "k" followed by another letter. Run the `table` command on the first character following the first "k" of each word.

```{r}
# I first find words that contain a "k" followed by another letter
d <- grep("[k].+[[:alpha:]]", a, value=TRUE)
# Then split words into a vector of letters
d1 <- strsplit(d, '')
# For vector d1, find where k is located within a word
d2 <- unlist(lapply(d1, function(x){y= x=='k'; return(which(y))}))

# Make a loop with the location of k, return the letter followed by k. 
d3 <- c()
for (i in 1:length(d2)){
  d3<- c( d3,d1[[i]][d2[i]+1])
}

# Make a table with d3.
table(d3)
```

8. Remove all vowels. How many character strings are found exactly once?

```{r}
e1 <- gsub("[aeiou]", '', a)
sum(table(e1)==1)
```
581 character strings are found once.


### Question 3 ###

**3 points**

The first argument to most functions that fit linear models are formulas.  The following example defines the response variable `death` and allows the model to incorporate all other variables as terms. `.` is used to mean all columns not otherwise in the formula.
l
```{r}
# I changed the dataset because error message came out
haart_df <- read.csv('~/downloads/haart.csv')[,c('death','weight','hemoglobin','cd4baseline')]
coef(summary(glm(death ~ ., data=haart_df, family=binomial(logit))))
```

Now imagine running the above several times, but with a different response and data set each time.  Here's a function:

```{r}
myfun <- function(dat, response) {
  form <- as.formula(response ~ .)
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}
```

Unfortunately, it doesn't work. `tryCatch` is "catching" the error so that this file can be knit to PDF.

```{r}
tryCatch(myfun(haart_df, death), error = function(e) e)
```


What do you think is going on?  Consider using `debug` to trace the problem.

The 'response' argument in the myfun function is not recognized as an object. So it doesn't work because R cannot recognize the variable such as death or hemoglobin, which makes it impossible to form a new function.


**5 bonus points**

Create a working function.
```{r}
myfun_new <- function(dat, response) {
  form <- as.formula( paste0(substitute(response), "~.")  )
  coef(summary(glm(form, data=dat, family=binomial(logit))))
}

myfun_new(haart_df, death)
```

So in this case, I included subsitute(response) so that the function could recognize the response variable name as a symbol for formula and use that symbol to create a formula.