---
title: 'Bios 6301: Assignment 2'
author: 'Jeongwon Choi'
output: html_document
---

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
```{r }
library(tidyverse)
library(dplyr)
setwd("C:/Users/choij33/OneDrive - Vanderbilt/Vanderbilt/Course/Statcomputing/Bios6301-main/Bios6301-main/datasets")

getwd()

cancer.df<-read.csv("cancer.csv")
```

    2. Determine the number of rows and columns in the data frame. (2)
    rows: 42120, cols: 8
```{r }
dim(cancer.df)
```
    3. Extract the names of the columns in `cancer.df`. (2)
```{r }
colnames_cancer.df<-colnames(cancer.df)
colnames_cancer.df
```

    4. Report the value of the 3000th row in column 6. (2)
```{r }
print(cancer.df[300, 6])
```

    5. Report the contents of the 172nd row. (2)
```{r }
print(cancer.df[172,])
```

    6. Create a new column that is the incidence *rate* (per 100,000) for each row. The incidence rate is the `(number of cases)/(population at risk)`, which in this case means `(number of cases)/(population at risk) * 100,000`. (3).
```{r }
cancer.df2<-cancer.df %>%
  mutate(rate=incidence/population * 100,000)
```


    7. How many subgroups (rows) have a zero incidence rate? (2)
    ```{r }
  sum(cancer.df2$rate==0)
    ```
    
 8. Find the subgroup with the highest incidence rate.(3)
```{r }
cancer.df2[cancer.df2$rate == max(cancer.df2$rate), ]
```
 
 
2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)

            max(x)
            sort(x)
            sum(x)
```{r}
x <- c("5","12","7")
max(x) #"7"
sort(x) # "12" "5"  "7"
#sum(x) #reason: character type cannot be numerically summed
```

    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12)
            y[2] + y[3]
```{r}
y <- c("5",7,12)
#y[2] + y[3] #It cannot be calculated because 7 and 12 are considered as character because "5" is a character (following the hierarchy of output type)
```
            

    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12)
            z[1,2] + z[1,3]
            
```{r}
z <- data.frame(z1="5",z2=7,z3=12)
z
z[1,2] + z[1,3] # It can be calculated because data.frame can hold different types of data - so z2 and z3 are double (not character) as they are.
```

3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
```{r}
c(1:8, 7:1)

```

    
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
```{r}
rep(1:5, 1:5)
```

    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
```{r}
d<-matrix(data=rep(1,9), nrow=3, ncol=3)
diag(d)<-rep(0, length(diag(d)))
d
```

    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
```{r}
d4 <- matrix(rep(1:4,4), nrow=4, ncol=4, byrow=TRUE)
for(i in 1:dim(d4)[1]){
  d4[i, ] <- d4[i,]^i
}
d4
```


4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. As an example, use `x = 5` and `n = 2`. (5 points)
```{r}
x<-5; n<-2
a <- 1
for(i in 1:n){
a <- a + x^i 
}
a
```

    1. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
```{r}
a1 <- 0
for (i in 1:999) {
  if (i %% 3 == 0 | i %% 5 == 0) {
    a1 <- a1 + i
  }
}
print(a1)
```

        1. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
```{r}
a2 <- 0
for (i in 1:999999) {
  if (i %% 4 == 0 | i %% 7 == 0) {
    a2 <- a2 + i
  }
}
print(a2)
```

    1. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])

```{r}


fib_seq <- c(1,1)
  for (i in 3:60) {
    fib_seq[i] <- fib_seq[i-1] + fib_seq[i-2]
  }
  
fib_seq

counter <-0; limit <- 15; summed <- 0
for (i in 1:length(fib_seq)){
  if (fib_seq[i]%%2==0){
    summed <- summed + fib_seq[i]
    counter <- counter + 1
  }
  if (counter >= limit){
    break
  }
}
summed
```

Some problems taken or inspired by projecteuler.

[euler1]: https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2
