---
title: 'Bios 6301: Assignment 8'
author: "Jeongwon Choi"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

*Due Tuesday, 15 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

30 points total.

**30/30, great work! See minor comments for Q2. Tianyi**

Submit a single knitr file (named `homework8.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework8.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

Install the `readxl` package and run the following

```{r}
fn <- 'icd10.xlsx'
if(file.access(fn, mode = 4) == -1) {
    url <- "https://www.cdc.gov/nhsn/xls/icd10-pcs-pcm-nhsn-opc.xlsx"
    download.file(url, destfile = fn, mode = 'wb')
}
dat <- readxl::read_excel(fn, sheet = 2)
```

1. Show the class of `dat`. (1 point)
```{r}
class(dat)
# attr(dat, "class")
```

2. Show the methods available for objects of the given class (if there are multiple classes, show methods for all classes). (3 points)
```{r}
methods( ,'tbl')
methods( ,'tbl_df')
methods( ,'data.frame')
```

3. If you call `print(dat)`, what print method is being dispatched? (1 point)
```{r}
print(dat)
```
There are 3 classes in dat - tbl, tbl_df, and data.frame. UseMethod searches for a match sequentially, for this case, tbl, tbl_df, and so on. If this cannot find an appropriate type of function, default function is used. So in this case, tbl is dispatched.

4. Set the class of `dat` to be a data.frame. (1 point)
```{r}
class(dat) <- c("data.frame")
#dat<- structure(list(), class = "data.frame")
class(dat)
```

5. If you call `print(dat)` again, what print method is being dispatched? (1 point)
```{r}
print.data.frame(dat)
print(dat)
```
In this case, print method of data.frame is being dispatched.

Define a new generic function `nUnique` with the code below.

```{r}
nUnique <- function(x) {
    UseMethod('nUnique')
}
```

6. Write a default method for `nUnique` to count the number of unique values in an element. (2 points)
```{r}
library(dplyr)
nUnique.default <- function(x, ...){
  n_distinct(x)
}
```

7. Check your function (2 points)

```{r, eval = FALSE}
nUnique(letters) # should return 26
nUnique(sample(10, 100, replace = TRUE)) # should return 10 (probably)
```

8. Write a data.frame method for `nUnique` to operate on data.frame objects.
This version should return counts for each column in a data.frame. (2 points)
```{r}
nUnique.data.frame <- function(x, ...){
  l1 <- lapply(x, unique)
  l2 <- lapply(l1, length)
  unlist(l2)
    }
```

9. Check your function (2 points)

```{r, eval = FALSE}
nUnique(dat)
```

### Question 2 ###

**15 points**

Programming with classes.  The following function will generate random patient information.

```{r}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class `medicalRecord` for objects that are a list with the named elements `name`, `gender`, `date_of_birth`, `date_of_admission`, `pulse`, `temperature`, `fluid_intake`. Note that an individual patient may have multiple measurements for some measurements.  Set the RNG seed to `8` and create a medical record by taking the output of `makePatient`.  Print the medical record, and print the class of the medical record. (5 points)
```{r}

# define class "medicalRecord"
medicalRecord <- setClass("medicalRecord", slots = c(name="character", gender="factor", date_of_birth="Date", date_of_admission="Date", pulse="numeric", temperature="numeric", fluid_intake="numeric"),contain="list")

# set random seed and generate a patient data
set.seed(8)
mr <- makePatient()
# names(mp) <- c('name', 'gender', 'date_of_birth', 'date_of_admission', 'pulse', 'temperature', 'fluid_intake')

# assign medicalRecord class to the data
mr.new <- new("medicalRecord",name=mr[[1]],gender=mr[[2]],date_of_birth=mr[[3]],
              date_of_admission=mr[[4]], pulse=mr[[5]], temperature=mr[[6]],
              fluid_intake=mr[[7]])

# show the class of the medical record
print(class(mr.new))

# print the medical record
print(mr.new)

```



2. Write a `medicalRecord` method for the generic function `mean`, which returns averages for pulse, temperature and fluids. Also write a `medicalRecord` method for `print`, which employs some nice formatting, perhaps arranging measurements by date, and `plot`, that generates a composite plot of measurements over time.  Call each function for the medical record created in part 1. (5 points)

```{r}
mean.medicalRecord <- function(x){
  return(data.frame(pulseavg=mean(x@pulse), tempavg=mean(x@temperature), fluiavg=mean(x@fluid_intake)))
}
```

```{r}
print.medicalRecord <- function(x){
  cat(paste("Name:", x@name, "\n", "Gender:", x@gender,"\n", "Date of birth:", x@date_of_birth,"\n"))
      
  cat("Measurement Date:\n")

  newd0 = data.frame(date_of_admission = x@date_of_admission, pulse = x@pulse, temperature = x@temperature, fluid_intake = x@fluid_intake)
  newd0 = newd0[order(newd0$date_of_admission), ]
  for(i in 1:length(newd0$date_of_admission)){
   print(paste0("Date: ", newd0$date_of_admission[i],  "|", "Pulse: ", newd0$pulse[i],"|", "Temp: ", newd0$temperature[i],"|", "Fluid_intake: ", newd0$fluid_intake[i]))
  }
}

# TS: check function
print(mr.new)
```

```{r}
plot.medicalRecord <- function(x){
  newd = data.frame(date_of_admission=x@date_of_admission, pulse=x@pulse, temperature=x@temperature, fluid_intake=x@fluid_intake)
  newd = newd[order(newd$date_of_admission), ]
 par(mfrow = c(1,3))
  plot(newd$date_of_admission, newd$pulse, type = "l", main = "pulse")
  # plot(newd$date_of_admission, newd$tempearture, type = "l", main = "temperature") #TS: newd$tempearture cannot be found and so it plots date_of_admission over index
    plot(newd$date_of_admission, newd$temperature, type = "l", main = "temperature")
  plot(newd$date_of_admission, newd$fluid_intake, type = "l", main = "fluid_intake")

}

plot(mr.new)
```


3. Create a further class for a cohort (group) of patients, and write methods for `mean` and `print` which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients.  Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for `mean` and `print`. (5 points)
```{r}

# define class "cohort"
cohort <- setClass("cohort",contain="list") # this is a just list consisting of medical records

# define methods for 'mean' and 'print'
mean.cohort = function(arr){
  
  df.cohort <- data.frame()
  
  for (i in 1:length(arr)){
    this_df <- data.frame(avg.pulse=mean(arr[[i]]@pulse), 
                          avg.temperature=mean(arr[[i]]@temperature),
                          avg.fluid_intake=mean(arr[[i]]@fluid_intake))
    df.cohort <- rbind(df.cohort, this_df)
  }
  return(df.cohort)
}
print.cohort = function(arr){
  
  for (i in 1:length(arr)){
    print(paste0('Patient ',i))
    print(arr[[i]]) # arr[[i]] has class of medicalRecord
  }
}

# set random seed
set.seed(8)

# generate a cohort of ten patients
patients <- lapply(1:10, function(d){return(makePatient())})

# transform each record into medicalRecord class
patients.mr <- lapply(patients,function(li){
  new("medicalRecord",name=li[[1]],gender=li[[2]],
date_of_birth=li[[3]],date_of_admission=li[[4]], pulse=li[[5]], temperature=li[[6]], fluid_intake=li[[7]])})

# assign cohort class to patient.mr
patients.cohort <- new('cohort',patients.mr)

# apply mean to cohort
# TS: It would be better if it shows patients' names (e.g. as rownames)
mean(patients.cohort)

# apply print to cohort
print(patients.cohort)

```

